Extended Abstract

### 2023-08-11 Analyses/extended_abstract_asa23.Rmd
edits 811

---
title: "Extended Abstract"
author: "Kevin Kiley & Nicolas Restrepo"
fontsize: 12pt
output:
  pdf_document: default
  html_document: default
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{lineno}
  - \linenumbers
  - \def\linenumberfont{\normalfont\tiny\sffamily}
---

@@

Adolescence and young adulthood are typically identified in the social science literature as "impressionable years" -- a stage of the life course in which people are more susceptible to broad social currents, when their opinions are more malleable, and when their core attitudes are formed []. Research regularly finds that people are more likely to change their attitudes and opinions in these life stages than other periods, with opinions becoming more stable once these stages have passed []. This pattern underpins dynamics around cohortization that are frequently invoked in larger models of social change [@ryder1965; @mannheim1952; @vaisey2016].

However, why people in these life stages appear to show more frequent opinion change relative to older people is unclear. Work in biology, psychology, and cultural evolution shows that peoples' cognition changes in ways that might make them more resistant to new information as they age []. Explanations in sociology, in contrast, tend to highlight the unique social features of these life stages -- like high rates of mobility between social contexts, and participation in educational institutions -- that shape how people are exposed to new sources of information or their ability to update opinions in the face of this information []. While both mechanisms likely play a role in facilitating opinion change in this life stage, the relative contributions of each are unclear.

Clarifying the mechanism that drives the frequency of opinion change during this formative period is important for understanding the broader process of social change. If openness to change in early adulthood is principally a function of the social structure of life in that time period, changes in the timing and ubiquity of life course events like marriage and childbearing can have significant impacts on the broader patterns of social change. Similarly, trends such as the deinstitutionalization of marriage [@cherlin2004; @cherlin2021; @brückner2005] that decrease distinctions between life course stages might make older people more open to change in more recent eras, diminishing the importance of "cohorts" for social change. If the mechanisms are principally developmental, rather than social, then these changes are unlikely to affect rates of social and cultural change in the same way.

Over the course of the 20th century, the timing of life-course transitions such as completing schooling, entering the workforce, getting married, leaving one's parents' home, and having children, have changed dramatically, with most of these event occurring later in the life course in more recent cohorts [@buchmann1989; @brückner2005].
Similarly, this stage of life has become more detached from family and other role obligations, giving people more freedom to explore new ideas and lifestyle choices [@rosenfeld2005]. This suggests both a lengthening and a intensification of this stage of the life course in ways that should increase the rates at which people update their attitudes. In this paper, we attempt to clarify the relative contributions of these different mechanisms by exploring how these social trends have affected rates of change in political attitudes over time, especially differences in rates of opinion change between young adults and the rest of the population.

## Hypotheses

Recent changes in the structuring of the life-course should help us adjudicate the relative contributions between social and developmental factors in rates of attitudinal updating among young adults. If their openness to information is purely developmental, then recent social changes should not affect rates of attitudinal change across young people from different periods (**H1**). However, if social structure plays a key role, we should be able to detect differential rates of change across these groups.

Sociological work outlines two specific trends that engender clear and testable predictions. The first is related to the elongation of young adulthood: people are increasingly postponing the rituals associated with "settling down". If social conditions affect the length of formative periods, then we should expect young people in more recent cohorts to exhibit higher rates of change (**H2**).

Nonetheless, work like that of Rosenfeld [] suggests that it is not only about the speed at which we move through the life-course, but also about the opportunities we have for exploration and self-expression. Contemporary social changes can be explained by the fact that people are more mobile and financially independent and, therefore, have more freedom to deviate from convention. This is an argument then about the *intensity* of formative periods. While this argument also predicts higher rates of absolute change among young people in more recent periods, it suggests that these differences will persists even when we account for differential timing of life events like marriage (**H3**).

@@

To test the hypotheses outlined above, we require data on within-individual change in comparable attitudes, over a stable window of time. We also need this data to capture different time periods within the same general population. To do this, we focus on general political beliefs, including views on the role of government, political ideology, and partisan identification, that have been asked in a variety of panel surveys. This is in part a practical matter, as these are the kinds of questions that have been included in panel studies that date back to the mid-1950s, giving us the largest window of cohorts to explore across comparable issues. However, these kinds of questions have frequently been used to identify the "impressionable years" [@alwin1991; @ghitza2022], so using these questions helps our results speak to these previous findings more directly.

We draw on data from four different survey panels: the 1956-60, 1972-76, and 1992-96 American National Election Study panels, and the 2006-10 General Social Survey Panels. All four surveys are nationally representative samples of non-institutionalized adults in the United States. All panels also survey respondents three times over a four-year window, making the duration comparable. They also include a variety of comparable questions about general sentiments toward government action, as well as measures of life course transitions.

The broadest challenge for our analysis is that it is not obvious what it means for a person to demonstrate opinion change in a survey. Thus, rather than adopt a single approach to quantify the rates of change across questions, we use a variety of approaches. We calculate whether participants changed at all across the three waves, we measure their absolute movement across the scales, whether they crossed the midpoint of the scale at any point, and whether their last poisition is at leasty two points away from their starting one. All these outcome variables presuppose different assumptions about what "change" entails and, thus, we deploy them all in an attempt to identify patterns that are robust across a range of assumptions.

For every definition of change, we regress the outcome measure on an indicator variable for whether people are aged 30 or less. For the continuous outcomes, we use ordinary least squares regression and, for the dichotomous outcomes, we use logistic regression. This produces an extremely large number of models for each of many questions, and will likely produce statistically significant coefficients at the $p < .05$ level for a substantial proportion of coefficients simply by chance.
Our goal is not to interpret any single coefficient but to interpret overall patterns if they emerge.

@@

@@

```{r, out.height='200%'}
model_paths <- list.files(path="../Models", pattern=".rds$", all.files=TRUE,
                          full.names=TRUE, recursive = T)

all_coefficients <- data.frame()

for (i in 1:length(model_paths)) {
  path <- model_paths[[i]]

  model <- read_rds(path)

  young_variance <- coef(model)$variable[, , 'young'] %>%
    as.data.frame()

  name <- str_remove(path, "../Models/")

  young_variance <- young_variance %>%
    mutate(model = substr(name, 1, 3),
           change_type = str_extract(name, "any|crossed|total|twopoints"),
           variable = rownames(young_variance))

  all_coefficients <- rbind(all_coefficients, young_variance)
}

all_coefficients %>% 
  filter(model %in% c("a5s", 
                      "a7s", 
                      "a9s", 
                      "gss")) %>% 
  mutate(model = case_when(
    model == "a5s" ~ "ANES56", 
    model == "a7s" ~ "ANES74", 
    model == "a9s" ~ "ANES94", 
    model == "gss" ~ "GSS06"
  )) %>% 
  ggplot(aes(x = variable, y = Estimate, col = model)) +
  geom_pointrange(aes(ymin = `Q2.5`, 
                      ymax = `Q97.5`), 
                  position = position_dodge(width = 0.8)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "darkred") +
  scale_color_brewer(palette="BrBG") +
  facet_wrap(~ change_type) + 
  theme(legend.position = "bottom") + 
  coord_flip() +
  labs(title = "Figure 1", 
       x = "", 
       y = "Predicted Difference")
```

@@

## Conclusion

>>>>

---
title: "Extended Abstract"
author: "Kevin Kiley & Nicolas Restrepo"
fontsize: 12pt
output:
  pdf_document: default
  html_document: default
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{lineno}
  - \linenumbers
  - \def\linenumberfont{\normalfont\tiny\sffamily}
bibliography: ./efw_bib.bib

---

@@

Adolescence and young adulthood are typically identified in the social science literature as "impressionable years" -- a stage of the life course in which people are more susceptible to broad social currents; when their opinions are more malleable; and when their core dispositions, identities, and opinions are formed [@alwin1991]. A broad body of research finds that rates of opinion change tend to peak during the window of ages 14-25, with opinions becoming more stable once people move into middle age [@sears1999; @bartels2014; @ghitza2022]. And this recurring empirical pattern is central to broader dynamics around cohortization that are frequently invoked in larger models of social change [@ryder1965; @mannheim1952; @vaisey2016].

However, why people in these life stages appear to show more frequent opinion change relative to older people is unclear. Work in biology, psychology, and cultural evolution suggests that peoples' cognition changes in ways that might make them more resistant to new information as they age. Explanations in sociology, in contrast, tend to focus on the unique social features of these life stages -- including high rates of mobility between social contexts and participation in educational institutions -- that shape how people are exposed to new sources of information and their ability to update opinions in the face of this information [@visser2004a]. While both mechanisms likely play a role in explaining the higher rates of opinion change in this life stage, the relative contributions of each are unclear.

In this paper, we leverage changes in the timing, ubiquity, and standardization of life course transitions over the second half of the 20th Century [@buchmann1989; @bruckner2005] to adjudicate the relative contributions of developmental and social mechanisms to the heightened rates of opinion change we observe in young people. Since the 1950s, the median age of marriage and childbirth have each been pushed back by about 10 years for both men and women [@eickmeyer2017; @manning2014], rates of college-going have increased dramatically, and people make more transitions between workplaces and between the workforce and education [@bruckner2005]. These shifts have led to a dramatically expanded window of participation in educational institutions, geographic and social mobility, and freedom from various social obligations and expectations, all of which are assumed to facilitate opinion change [@rosenfeld2009]. If the heightened rates of opinion change we observe in young adults are principally due to the social structure of this life stage, then we should observe changes in the rates of opinion change in this group over time.

Drawing on four nationally representative panel surveys, dating from the mid-1950s through the early 2000s, we compare multiple different measures of opinion change in general political opinions and partisan identification of early adults (people under 30) and the rest of the population. Despite the significant changes in the standardization of the life course and the duration of early adulthood, we find remarkable consistency in opinion behaviors over time, both for young adults and for the rest of the population. Across multiple measures of opinion change, we find that people under 30 change their opinions at slightly higher rates than people over 30, but this gap remains stable across panels. The overall pattern of results suggests either that the heightened rates of attitude change in early adulthood are principally driven by relatively constant developmental factors -- how the brain reacts to new information as people age -- or that the dramatic changes in the social structuring of early adulthood have combined in a way to produce aggregate stability over six decades.

@@

We draw on data from four different panel surveys of the American population: the 1956-60, 1972-76, and 1992-96 American National Election Study panels, and the 2006-10 General Social Survey panel. All four surveys are nationally representative samples of non-institutionalized adults in the United States. All panels also interviewed survey respondents three times over a four-year window, making the duration of each survey window comparable. All four panels also include comparable questions about general sentiments toward government action, a measure of partisan identification, and measures of life course transitions such as marriage and childbearing.

The broadest challenge for our analysis is that it is not obvious what it means for a person to demonstrate opinion change in a survey. Thus, rather than adopt a single approach to quantify the rates of change across questions, we use a variety of approaches. We calculate whether participants changed at all across the three waves (), their absolute movement across the scales (), whether they crossed the midpoint of the scale at any point (), and whether their last position is at least two points away from their starting one (). All these outcome variables presuppose different assumptions about what "change" entails and, thus, we present them all in an attempt to identify patterns that are robust across a range of assumptions.

For every definition of change, we regress the outcome measure on an indicator variable for whether people are aged 30 or less. For the continuous outcomes, we use ordinary least squares regression and for the dichotomous outcomes we use logistic regression. This produces an extremely large number of models for each of many questions, and will likely produce statistically significant coefficients at the $p < .05$ level for a substantial proportion of coefficients simply by chance. Our goal is not to interpret any single coefficient but to interpret overall patterns if they emerge.

@@

Figure 1 plots coefficient estimates generated from regression models of change on whether respondents are under 30. Each column represents a different measure of "change" outlined above. ... If young people have become more likely to make

```{r}

```

@@

@@

This is a rather striking pattern that challenges the notion that

## Discussion
